### 2016-09-29 rebuttal.tex

\begin{verbatim}
   >>> phi = (1+sqrt(5))/2
   >>> psi = (1-sqrt(5))/2
   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
   65536.0000000000
   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
   0.e-104
\end{verbatim}

Obviously, like your own example, these are problematic because a part of the computation is relying on Python's builtin floats. Please include a comment on whether symbolic computation (i.e., applying \texttt{evalf()} on the whole expression) always avoids these errors or not, possibly with an example when it doesn't resolve this problem.
\begin{solution}
  Your first example, unfortunately, doesn't work due to automatic
  simplification (see the last input line above:  \texttt{exp(100)} cancels
  automatically, before any numeric evaluation happens). Your second example
  doesn't have this issue, but our example is simpler.

Numerical evaluation of the whole expression should
always solve cancellation problem, as it was stated in the sentence
``Applying the evalf method to the whole expression solves this problem.''
\end{solution}

\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
\begin{solution}
The mpmath version is now mentioned in the introduction.
\end{solution}

@@

\question In line 495, the word ``solving'' seems more appropriate than ``solutions''.
\begin{solution}
This was replaced with ``solving ODEs''.
\end{solution}

\question In line 504 ``is'' should be used instead of ``are'' (because ``array'' is singular).
\begin{solution}
This was fixed.
\end{solution}

\question In line 518, ``produces'' should become plural.
\begin{solution}
This was fixed.
\end{solution}

\question The title of section 5, ``Domain specific submodules'', seems inappropriate because the section only covers Physics package (not ``submodules''). It should either be expanded with a short introduction listing other domain specific packages, or it should be renamed to ``Physics package''.
\begin{solution}
\label{rev3point30}
This section (section~\ref{P-sec:domain_specific}) was renamed to ``Physics submodule''.  (See also the
discussion at~\ref{rev3point37} about module vs package).
\end{solution}

### 2016-09-29 rebuttal.tex

\section{Reviewer 1}
\subsection{Basic reporting}
\begin{questions}
\question Generally good. A part that confused me is the assertion (footnote 3) that ``If
$A$ and $B$ are Symbols created with \texttt{commutative=False} then SymPy will keep
$A\cdot B$
and $B\cdot A$ distinct.'' Does that mean that BOTH of them must be created this way,
and that $A$ and $x$ (if $x$ is created normally) will commute? Is there any way to
declare commutators? How does one guarantee that other pieces of code, e.g.
Gaussian elimination, respect non-commutativity?
\begin{solution}
  We have clarified the footnote. Both expressions must be set as
  \texttt{commutative=False}. Internally, in \texttt{Mul}, the ``commutative
  part'' of an expression is pulled out to the front and canonically ordered,
  and the ``noncommutative part'' is not reordered.

Commutators are implemented in the \texttt{sympy.physics.quantum} submodule and
are discussed in section~\ref{P-sec:quantum}.

@@

We disagree with the reviewer on a few points. We do not consider SymPy Live
or SymPy Gamma to be key components of SymPy. The goal of the paper is
to be about the SymPy library. While SymPy Live and SymPy Gamma are
maintained by the SymPy community, they are separate projects. A
mention of SymPy Live has been added to the introduction (see editor
point~\ref{editorpoint1}), as it may assist an interested reader in trying the
examples from the paper. A mention of SymPy Gamma has been added to the ``Projects that
Depends on SymPy'' table (Table~\ref{P-projects-table}), now part of the main
paper (see above). We have
left the more in depth discussion of SymPy Gamma in the supplement (section~\ref{S-suppsec:sympy-gamma}).

The ``Comparison with Mathematica'' section has been placed in the supplement
because we do not feel that a treatment in the main manuscript would be fair
without an equal comparison to other principle computer algebra systems, such
as Maple, SageMath, and Maxima. Unfortunately, our authorship is inexpert in
these systems, so a complete and fair comparison is impossible.

We deem all the sections in section 3 (now section~\ref{P-sec:features}) to be important core
components of the library. Regarding the supplement sections (previously) 5,
7, and 9---``Sets'', ``Category Theory'', and ``SymPy Live'',
respectively:
\begin{itemize}
\item In SymPy 1.0, the sets submodule is still in
a relatively infant stage. It is mentioned already in conjunction with the
\texttt{solveset} function. We have added some additional text to the solvers
subsection regarding \texttt{sympy.sets}.

\item The category theory submodule in SymPy
is very domain specific, and independent of other SymPy submodules. It is out of
scope for an architecture paper.

@@

\item The SymPy Live section was short and would be mostly duplicated by the
  paragraph that has been added to the introduction in the paper. We have thus
  removed it.
\end{itemize}

We also consider the supplement sections (previously) 2, 3, 4, and
6---``Series'', ``Logic'', ``Diophantine Equations'', and ``Statistics'',
respectively---to be in depth beyond the goals of the main manuscript:

\begin{itemize}
\item The ``Series'' section discusses relatively advanced series expansion methods. We
have added a paragraph detailing basic series expansion to the ``Calculus''
subsection, with a reference to the more detailed section in the supplement.

\item The ``Logic'' section is alluded to in the ``Assumptions'' subsection,
  and the submodule is briefly mentioned in the ``Extensibility'' section,
  but is not discussed in depth in the main manuscript. While the logic
  submodule is indeed important for certain parts of SymPy, we do not consider
  it to be of interest to the general reader. Furthermore, on a technical
  note, the assumptions discussed in the manuscript are the so-called ``old
  assumptions'', which do not use the \texttt{sympy.logic} submodule discussed
  in the supplement (they use an implementation of the Rete algorithm, which
  is separate from \texttt{sympy.logic}). The so-called ``new assumptions'',
  mentioned briefly in the ``Conclusion and Future Work'' section, does use
  it, but we have opted to not discuss this submodule in depth, as it is still in
  development and not fully usable. We have added some text to the assumptions
  section discussing this and referencing the logic subsection of the
  supplement (see footnote~\ref{P-note:assumptions}).

@@

\question In lines 567--568, ``performing symbolic quantum mechanics'' makes no sense. This should probably be ``computations'', ``solving problems related to'', etc.
\begin{solution}
This was reworded to:
\begin{quote}
to solve problems in quantum mechanics
\end{quote}
\end{solution}

\question The sentence ``SymPy expressions are immutable trees of Python objects.'' doesn't belong in the conclusion. This can be moved to the appropriate place when discussing SymPy's architecture.
\begin{solution}
This paper focuses on the architecture of SymPy, and we consider this to be a
core feature of SymPy's architecture, hence, it's inclusion in the conclusion.
We have also discussed the different points in this sentence in depth in the architecture section.
\end{solution}

\question All ``submodules'' should be replaced by ``modules'' (examples:
lines 657, 662).
\label{rev3point37}
\begin{solution}
  Our use of ``modules'', ``submodules'', and ``packages'' coincides with the
  definitions used in the official Python documentation (see
  \url{https://docs.python.org/3/tutorial/modules.html#packages} and
  \url{https://docs.python.org/3/glossary.html#term-package}). To clarify
  things, we have modified the manuscript to use ``package'' only when
  referring to a library external to SymPy (such as mpmath and Xy-pic).

To avoid confusion, and to be consistent with the definitions used by the
official Python documentation, we have modified the manuscript to use
``submodule'' to refer to any module that is a dotted submodule of
\texttt{sympy}, such as \texttt{sympy.physics.quantum}. This is to emphasize
that these submodules all live within the \texttt{sympy} namespace, as ``module''
can suggest a separate library.
\end{solution}

### 2016-09-29 conclusion_and_future_work.tex
CASs -> CASes
Reviewer 3, comment 3.
Match the conclusion to the current paper sections

SymPy is a robust computer algebra system that provides a wide spectrum of
features both in traditional computer algebra and in a plethora of scientific
disciplines. This allows SymPy to be used in a first-class way with other
Python projects, including the scientific Python stack. Unlike many other CASs, SymPy
is designed to be used in an extensible way: both as an end-user
application and as a library.

@@

SymPy supports a wide array of mathematical facilities. This includes functions for
simplifying expressions, performing common calculus operations, pretty printing
expressions, solving equations, and representing symbolic matrices. Other supported
facilities
include discrete math, concrete math, plotting, geometry, statistics,
polynomials, sets, series, vectors, combinatorics, group theory, code
generation, tensors, Lie algebras, cryptography, and special functions.
Additionally, SymPy contains submodules targeting certain specific domains,
such as classical mechanics and quantum mechanics.  This breadth of domains has
been engendered by a strong and vibrant user community.
Anecdotally, these users likely chose SymPy because of its ease of access.

>>>>

SymPy is a robust computer algebra system that provides a wide spectrum of
features both in traditional computer algebra and in a plethora of scientific
disciplines. This allows SymPy to be used in a first-class way with other
Python projects, including the scientific Python stack. Unlike many other CAS's, SymPy
is designed to be used in an extensible way: both as an end-user
application and as a library.

@@

SymPy supports a wide array of mathematical facilities. This includes
functions for assuming and deducing common mathematical facts,
simplifying expressions, performing common calculus operations, polynomial
manipulations, pretty printing
expressions, solving equations, and representing symbolic matrices. Other supported
facilities
include discrete math, concrete math, plotting, geometry, statistics,
polynomials, sets, series, vectors, combinatorics, group theory, code
generation, tensors, Lie algebras, cryptography, and special functions.
Additionally, SymPy contains submodules targeting certain specific physics domains,
such as classical mechanics and quantum mechanics.  This breadth of domains has
been engendered by a strong and vibrant user community.
Anecdotally, these users likely chose SymPy because of its ease of access.
